Synthetic Observations of Extremely High Velocity (EHV) SiO components in young molecular outflows

### 2014-01-22 PAPER/Paola_2/ms_ver3.tex
Modified Figures and its captions II

\begin{document}
%\title{SiO Molecular Jets around young stars - A numerical perspective}
\title[EHV SiO components in Molecular Outflows]{Synthetic Observations of Extremely High Velocity (EHV) SiO components
  in young molecular outflows}
\author[B. Vaidya, T. Douglas, P. Caselli]{Bhargav Vaidya$^{1}$\thanks{E-mail:
B.Vaidya@leeds.ac.uk (BV)}, Thomas Douglas$^{1}$, Paola Caselli$^{1}$\\
$^{1}$School of Physics and Astronomy, University of Leeds, Leeds LS2
9JT\\
}

@@

\section{Introduction}
Jets are one of the first manifestations of star
formation in dense molecular cores. They are ubiquitous in both
massive and low-mass star forming regions. These supersonic flows,
perpendicular to the underlying accretion disk, play a vital role in
removing excess angular momentum and thereby aiding in the accretion onto the protostar. 
For low-mass stars, they are believed to be launched by magneto-centrifugally forces and
further collimated by magnetic hoop stress
\citep[][]{Blandford:1982p892, Konigl:2000p607}. However, in case of
high-mass stars, radiative forces also contribute to flow dynamics
during the later evolutionary stages (\citealt{Vaidya:2011p8992}). 
Typically, jets are a few parsec in size and can be divided into
three length-scale domains, viz. source and disk scales (1-10$^{2}\,$AU),
envelope scales ($10^{2} - 10^{5}$\,AU) and parent cloud scales
($10^{5} - 10^{6}$\,AU) (\refer : Frank et. al, 2013, PPVI).
Of these, it is at the envelope scale where rich
chemical evolution occurs as the jet interacts
with the molecular medium. In this region, the jet propagates into a
relatively static medium inducing shocks that are interesting from both a
physical and chemical point of view. In addition to jets, molecular
material from the surroundings is entrained and accelerated to high
velocities giving rise to molecular outflows.
%

@@

\subsection{Cooling Prescriptions}
%
Figure~\ref{fig:tempvar} shows the variation
of the cooling function, $\Lambda(\rho,T,{\bf{X}})$, with gas
temperature assuming conditions of initial ambient density. 
Various processes that contribute to the cooling function in our model
are listed in the figure. The formulations of these contributions
are adopted from \citep[][and references therein]{Smith:2003p9985, OSullivan:2009p10078}.
For low temperatures, we also 
take into account contributions to the cooling function from molecules like
CO, whose abundance with respect to $n_{\rm H}$ is fixed to a
value of 10$^{-5}$ assuming an average CO depletion factor across the
envelope of $\sim$ 10 \citep[e.g.][]{Crapsi:2005p16655}. 
For each of the other molecules contributing to the cooling function (like OH, H$_{2}$O and atomic OI)
the abundance is fixed to values of $5\times10^{-6}$. Note, the
uncertain abundance values are rather arbitrary for these molecules. 
For example, the adopted abundance of H$_{2}$O is higher than that found by
\citealt{Caselli:2012p16713} in a pre-stellar core ($\lesssim$
10$^{-7}$). Inspite of such over estimated abundance values, their
contribution is only important when temperature is $<$100\,K as seen in
figure~\ref{fig:tempvar}. As the internal shocks, where most
of the molecular hydrogen is produced, maintain temperatures higher
than 100\,K and densities $\sim$10$^{6}$ cm$^{-3}$, cooling due to these molecules do not
play a significant role even if the arbitrary abundance is increased
by a factor of 10 \citep{Smith:2003p9985}. Instead, a substantial contribution to the cooling
function, between T = 100\,K and 3000\,K, 
comes from the vibrational and rotational transitions of
molecular hydrogen during the initial stages of jet
evolution. For temperatures above 3000\,K the
cooling is mainly via H$_{2}$. The total curve for the
whole temperature range obtained in our runs is shown in black.
The fact that the initial ambient medium is largely molecular, cooling is also dominated by
molecular processes. In our reference model (see ~\ref{ssec:refrun}), 
we update this curve at each time step based on the fraction of
various hydrogen species. Thus, during the evolution, 
contributions from collisional ionizations and excitations become
important in regions dominated by atomic and ionized hydrogen.
%tabulated

@@

\begin{table*}
\centering 
\caption{Summary of parameter runs for different cooling prescriptions and density contrasts using two abundance profiles.}
\begin{tabular}{c | c | c | c | c | c | c }
\hline
Run & Cooling Mode & $\eta$ & \multicolumn{2}{|c|}{Top Hat Profile} & \multicolumn{2}{|c|}{Gaussian Profile}\\
\hline\hline
&&& $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] & $\int T_{\rm MB}dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] \\
\hline
adi10 & Nil (Adiabatic) & 10 & 2.91 & $>$40 & 0.25 & 40.0 \\
pow10 & Power law & 10 & 0.64 & 8.0 & 0.02 & 10.0\\
atm10 & Atomic & 10 & 2.04 & 36.0 & 0.58 & 38.0 \\
atm2 & Atomic & 2 & 3.21 & 18.0 & 0.64 & 20.0 \\
tab10 & Tabulated & 10 & 0.75 & 11.0 & 0.09 & 10.0 \\
tab2 & Tabulated & 2 & 2.89 & 8.0 & 1.0 & 9.0 \\
mol10 & Molecular & 10 & 1.10 & 10.0 & 0.1 & 11.0 \\
mol3 & Molecular & 3 & 3.3 & 14.0 & 0.66 & 12.0\\
\hline
\end{tabular}
\label{tab:result1}
\end{table*}

@@

The dynamical states for each of this cooling prescriptions are then 
given as inputs to the radiative transfer code to obtain integrated SiO(2-1) emission
maps shown in figure~\ref{fig:coolcmpsio}. The differences in dynamical
features are also reflected in the corresponding SiO emission
maps. These maps are obtained assuming an angle of inclination of
90$^\circ$ using the top hat abundance
profile. They are then placed at a distance of
16.7$\,$pc and convolved with a beam with FWHM of 2.0\arcsec, in order
to create the maps shown. 
The emission from the jet with atomic
cooling is coming from the conical jet head and younger knots closer
to base of the flow. Additionally, features in the shell at Z/R$_{\rm jet}$ $\sim$
60, 75 also show evidence of faint emission. In case of the jet with
power-law cooling, the emission mostly arises from the
internal knots. 
The emission coming from jets with
tabulated and molecular cooling resembles that observed in young
molecular outflows, although the tabulated jet is narrower.
Also in these outflows, strong emission arises from the
condensation formed at the bow shock due to thermal
instability. 
%

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig5.pdf}%diffcool_img.pdf}
 \caption{A plot of the integrated SiO(2-1) emission from 5 models, 
each using a different method to calculate cooling (labelled on the
top of each panel) and all with $\eta$=10.}
\label{fig:coolcmpsio} 
\end{figure*}

@@

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/pfig6.pdf}%hfracs_ref.pdf}
 \caption{{\it{Left}}: Colour composite image of various hydrogen
   fractions, fHI (red), fH$_{2}$ (green) and fHII (blue), for the
   reference run. For better visualisation, the ionized hydrogen
   fraction is enhanced by a factor of 100. The temperature of the jet
 obtained from dynamical simulation is shown on the {\it{right}}.}
\label{fig:Hfracsim}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/pfig8.pdf}%emspecpv_90.pdf}%{\figpath/refrun_21_90_emspecpv.png}
 \caption{Integrated emission map ({\it top left}), PV diagram ({\it top right}
   and spectra ({\it bottom}) for the
   SiO(2-1) transition for the reference run. 
   The jet is assumed to be in the plane of sky implying an angle of
   inclination of $90^{\circ}$. The vertical {\it magenta dashed} line
   represents the cut for the PV digram and {\it green} crosses marks the positions from which the spectra are taken.} 
\label{empvspec90}
\end{figure}

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig10.pdf}%diffabun_img.pdf}%{\figpath/EM_gaussandth.png}
 \caption{Integrated intensity maps of SiO(2-1) line in units of K, for runs with molecular
   cooling and having different abundance profiles.}
\label{fig:diffabunem}
\end{figure*}

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig13.pdf}%imshkvhcontomaps.png}
 \caption{Symmetrical contour maps of multi-line integrated SiO
   emission convolved with a 2\arcsec beam obtained using parameters of the reference run. The contour colors represent different intensities in Kelvins, i.e, 50.0({\it red}), 30.0({\it green}), 20.0({\it blue}), 5.0({\it
     magenta}), 2.0({\it cyan}), 0.1({\it black}).} 
\label{fig:ltconts}
\end{figure*}

\cite{Nisini:2007p13128} have shown that the current
plane-parallel shock models fit reasonably well the physical
conditions traced by the SiO emission i.e., temperature $\lesssim$
1000\,K and H$_{2}$ number density $\sim$ 10$^{5}$-10$^{6}$
cm$^{-3}$, assuming optically thin emission. 
They can also fit the observed fractional SiO abundance of
$\sim$ 10$^{-7}$ in these outflows. However, {\em they fail to predict all
the line profiles and in particular their similarities for low and
high excitation lines, which we are able to obtain.} 
Line emission for SiO from our model also traces similar
physical environment. Fig.~\ref{fig:Hmolform} shows that regions close
to the internal knot have temperatures up to 1000\,K and molecular
hydrogen density of the order of 10$^{6}$ cm$^{-3}$ (also see
fig.~\ref{grid}). These are the regions where the bulk of the SiO
emission is present, with fractional abundances between 10$^{-6}$ to 10$^{-8}$.
The line profiles shapes obtained from our study fit reasonably well that observed by
\cite{Nisini:2007p13128}, especially for the bullets seen in
L1448. {\em Interestingly, in our models, the line profiles are equally broadened for low as well
as high excitation lines}. The top panel of fig.~\ref{fig:lineratio} shows spectra for three line
transitions for the reference run but with an angle of inclination
$i$ = 60$^{\circ}$ and convolved with a single dish beam of
15\arcsec, to simulate observed integrated
intensities of L1448 bullets observed with large beams from JCMT and
IRAM. Intensities predicted from our reference model, $\int T_{\rm MB}
dV$ = 5.76 K\,km\,s$^{-1}$ for SiO(2-1), 
lie within a factor of two from the observed values (Table 2 of \citealt{Nisini:2007p13128}).
%

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig14.pdf}%singledish_new.pdf}%{\figpath/LineRatio_THDeq1.png}
 \caption{{\em{Top}}: Line profiles in SiO (2-1), (5-4) and (8-7) at one
   of the inner knots for the reference molecular cooling run. The profiles are
   obtained when the angle of inclination is 60$^{\circ}$ with respect
 to line of sight. {\em{Bottom left}}: Line intensity ratios SiO(8-7)/(5-4)
and SiO(5-4)/(2-1), as a function of velocity. {\em{Bottom right}}:
Variation of normalised integrated intensity with upper line transition J$_{\rm
  up}$ for the reference run.}
\label{fig:lineratio}
\end{figure*}

\subsection{Slow Component and Grain Chemistry}
%
We have presented a simple model that is able to explain the spectral characteristics
of the SiO EHV component seen
in several young outflows. The reference model
consistently solves for the H$_{2}$ chemistry in the presence of
appropriate cooling. The temperature, density and velocities obtained from dynamical
simulation is used as input in the non-LTE radiative transfer code to obtain SiO maps, spectra and PV
diagrams. One essential ingredient required for the radiative transfer modeling
is the SiO fractional abundance and its dependence on shock velocity. 
Though 1D models that study the formation of SiO in the gas phase from
sputtering and grain-grain 
collisions exists, their focus is mainly on weaker shocks
\citep{Schilke:1997p14140, Caselli:1997p14853, Gusdorf:2008p13800, VanLoo:2013p16573}. 
The pulsating jet propagation model presented here have bulk velocities
reaching up to 120 km\,s$^{-1}$. We have used few
simplified SiO abundance profiles as a function of jet
velocity. Though the shape of the profile is uncertain, the upper and
lower limits used are backed by observational evidences. Ideally, for a consistent
dependence of SiO abundance on shock velocity, one would have to also
solve for silicon and SiO chemistry in a manner similar to that of
H$_{2}$. However, such a numerically expensive model 
with complex chemistry is beyond the scope of this paper 
as this could imply to include grain processing and more gas-phase
chemistry in an MHD multi-fluid approach.  
%

@@

\section{Treatment of Adiabatic Index $\gamma$}
\label{sec:treatgamma}
In dealing with a gas chemistry network which can have both monatomic and diatomic
species, one is immediately presented with the issue of how to treat the adiabatic
index, $\gamma$, of the gas, which directly influences the equation of
state, eq.~\ref{EOS}, and thus the behaviour of the gas.
In standard hydrodynamical codes, the gas is generally
assumed to be a monatomic ideal gas, with $\gamma$ = 5/3. 
A diatomic gas, such as H$_{2}$, on the other hand, has a value of $\gamma$
= 1.4 at low temperatures, decreasing to below 1.3
at higher temperatures as the vibrational modes of the hydrogen molecule become
excited. Thus, the value of $\gamma$ for a gas containing a mixture of atomic and
molecular hydrogen requires a temperature-dependent calculation of the $\gamma$
value for H$_2$ and a weighting of the values for the atomic and molecular components
according to their abundance. \cite{Yoshida:2006p13376} have suggested
a prescription of an effective $\gamma$ which takes into account the
corresponding weights of each chemical species and is given by,
\begin{equation}
\frac{1}{\gamma_{\rm eff} - 1} = \frac{1}{\mu}\left[\sum_{i}\frac{X_{i}}{A_{i}}\frac{1}{(\gamma_{i} - 1)}\right]^{-1},
\label{eq:gammaeff}
\end{equation}
where i is the index for each specie and $\mu$ is the mean molecular
weight. The mass fraction is added in the above equation following the
correction suggested in \cite{OSullivan:2009p10078}.
Additionally, the temperature dependence in the case of
H$_{2}$ molecule can be obtained using \cite{landau:1980},
\begin{equation}
\frac{1}{\gamma_{H_{2}} - 1} = \frac{1}{2}\left[ 5 +
  2x^{2}\frac{e^{x}}{(e^{x} - 1)^{2}}\right],
\label{eq:gammaH2}
\end{equation}
where the last term with  x = 6100 $K$/T accounts for the vibrational
degree of freedom, with T being the gas temperature.
%

>>>>

\begin{document}
%\title{SiO Molecular Jets around young stars - A numerical perspective}
\title[EHV SiO components in Molecular Outflows]{Synthetic Observations of Extremely High Velocity (EHV) SiO components
  in young molecular outflows}
\author[B. Vaidya, T. Douglas, P. Caselli]{B. Vaidya$^{1}$\thanks{E-mail:
B.Vaidya@leeds.ac.uk (BV)}, T. A. Douglas$^{1}$, P. Caselli$^{1}$\\
$^{1}$School of Physics and Astronomy, University of Leeds, Leeds LS2
9JT\\
}

@@

\section{Introduction}
Jets are one of the first manifestations of star
formation in dense molecular cores. They are ubiquitous in both
massive and low-mass star forming regions. These supersonic flows,
perpendicular to the underlying accretion disk, play a vital role in
removing excess angular momentum and thereby aiding in the accretion onto the protostar. 
For low-mass stars, they are believed to be launched by magneto-centrifugally forces and
further collimated by magnetic hoop stress
\citep[][]{Blandford:1982p892, Konigl:2000p607}. However, in case of
high-mass stars, radiative forces also contribute to flow dynamics
during the later evolutionary stages (\citealt{Vaidya:2011p8992}). 
Typically, jets are a few parsec in size and can be divided into
three length-scale domains, viz. source and disk scales (1-10$^{2}\,$AU),
envelope scales ($10^{2} - 10^{5}$\,AU) and parent cloud scales
($10^{5} - 10^{6}$\,AU) \citep[e.g.,][and references therein]{Ray:2007p3306}.
Of these, it is at the envelope scale where rich
chemical evolution occurs as the jet interacts
with the molecular medium. In this region, the jet propagates into a
relatively static medium inducing shocks that are interesting from both a
physical and chemical point of view. In addition to jets, molecular
material from the surroundings is entrained and accelerated to high
velocities giving rise to molecular outflows.
%

@@

\subsection{Cooling Prescriptions}
%
Figure~\ref{fig:tempvar} shows the variation
of the cooling function, $\Lambda(\rho,T,{\bf{X}})$, with gas
temperature assuming conditions of initial ambient density. 
Various processes that contribute to the cooling function in our model
are listed in the figure. The formulations of these contributions
are adopted from \citep[][and references therein]{Smith:2003p9985, OSullivan:2009p10078}.
For low temperatures, we also 
take into account contributions to the cooling function from molecules like
CO, whose abundance with respect to $n_{\rm H}$ is fixed to a
value of 10$^{-5}$ assuming an average CO depletion factor across the
envelope of $\sim$ 10 \citep[e.g.][]{Crapsi:2005p16655}. 
For each of the other molecules contributing to the cooling function (like OH, H$_{2}$O and atomic OI)
the abundance is fixed to values of $5\times10^{-6}$. Note, the
uncertain abundance values are rather arbitrary for these molecules. 
For example, the adopted abundance of H$_{2}$O is higher than that found by
\citealt{Caselli:2012p16713} in a pre-stellar core ($\lesssim$
10$^{-7}$). In-spite of such over estimated abundance values, their
contribution is only important when temperature is $<$100\,K as seen in
figure~\ref{fig:tempvar}. As the internal shocks, where most
of the molecular hydrogen is produced, maintain temperatures higher
than 100\,K and densities $\sim$10$^{6}$ cm$^{-3}$, cooling due to these molecules do not
play a significant role even if the arbitrary abundance is increased
by a factor of 10 \citep{Smith:2003p9985}. Instead, a substantial contribution to the cooling
function, between T = 100\,K and 3000\,K, 
comes from the vibrational and rotational transitions of
molecular hydrogen during the initial stages of jet
evolution. For temperatures above 3000\,K the
cooling is mainly via H$_{2}$. The total curve for the
whole temperature range obtained in our runs is shown in black.
The fact that the initial ambient medium is largely molecular, cooling is also dominated by
molecular processes. In our reference model (see ~\ref{ssec:refrun}), 
we update this curve at each time step based on the fraction of
various hydrogen species. Thus, during the evolution, 
contributions from collisional ionizations and excitations become
important in regions dominated by atomic and ionized hydrogen.
%tabulated

@@

\begin{table*}
\centering 
\caption{Summary of parameter runs for different cooling prescriptions
  and density contrasts. Integrated
  emission map for SiO(2-1) line in units of K-km\,s$^{-1}$ obtained
  by convolving with a beam of 15\arcsec\,and the
  corresponding line widths (in units of km\,s$^{-1}$) are listed for each of these runs for two different abundance profiles.}
\begin{tabular}{c | c | c | c | c | c | c }
\hline
Run & Cooling Mode & $\eta$ & \multicolumn{2}{|c|}{Top Hat Profile} & \multicolumn{2}{|c|}{Gaussian Profile}\\
\hline\hline
&&& $\int T dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] & $\int T dV$ [K-km\,s$^{-1}$] & $\Delta$v [km s$^{-1}$] \\
\hline
adi10 & Nil (Adiabatic) & 10 & 2.91 & $>$40 & 0.25 & 40.0 \\
pow10 & Power law & 10 & 0.64 & 8.0 & 0.02 & 10.0\\
atm10 & Atomic & 10 & 2.04 & 36.0 & 0.58 & 38.0 \\
atm2 & Atomic & 2 & 3.21 & 18.0 & 0.64 & 20.0 \\
tab10 & Tabulated & 10 & 0.75 & 11.0 & 0.09 & 10.0 \\
tab2 & Tabulated & 2 & 2.89 & 8.0 & 1.0 & 9.0 \\
mol10 & Molecular & 10 & 1.10 & 10.0 & 0.1 & 11.0 \\
mol3 & Molecular & 3 & 3.3 & 14.0 & 0.66 & 12.0\\
\hline
\end{tabular}
\label{tab:result1}
\end{table*}

@@

The dynamical states for each of this cooling prescriptions are then 
given as inputs to the radiative transfer code to obtain integrated SiO(2-1) emission
maps shown in figure~\ref{fig:coolcmpsio}. The differences in dynamical
features are also reflected in the corresponding SiO emission
maps. These maps are obtained assuming an angle of inclination of
90$^\circ$ using the top hat abundance
profile. They are then placed at a distance of
16.7$\,$pc and convolved with a beam with FWHM of 2.0\arcsec\,in order
to create the maps shown. 
The emission from the jet with atomic
cooling is coming from the conical jet head and younger knots closer
to base of the flow. Additionally, features in the shell at Z/R$_{\rm jet}$ $\sim$
60, 75 also show evidence of faint emission. In case of the jet with
power-law cooling, the emission mostly arises from the
internal knots. 
The emission coming from jets with
tabulated and molecular cooling resembles that observed in young
molecular outflows, although the tabulated jet is narrower.
Also in these outflows, strong emission arises from the
condensation formed at the bow shock due to thermal
instability. 
%

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig5.pdf}%diffcool_img.pdf}
 \caption{A plot of the integrated SiO(2-1) emission from 5 models in
   units of K-km\,s$^{-1}$, 
each using a different method to calculate cooling (labelled on the
top of each panel) and all with $\eta$=10.}
\label{fig:coolcmpsio} 
\end{figure*}

@@

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/pfig6.pdf}%hfracs_ref.pdf}
 \caption{{\it{Left}}: Color composite image of various hydrogen
   fractions, fHI (red), fH$_{2}$ (green) and fHII (blue), for the
   reference run. For better visualisation, the ionized hydrogen
   fraction is enhanced by a factor of 100. The temperature of the jet
 obtained from dynamical simulation is shown on the {\it{right}}.}
\label{fig:Hfracsim}
\end{figure}

@@

\begin{figure}
 \includegraphics[width=1.\columnwidth]{\figpath/pfig8.pdf}%emspecpv_90.pdf}%{\figpath/refrun_21_90_emspecpv.png}
 \caption{Integrated emission map ({\it top left}) in units of K-km\,s$^{-1}$, PV diagram ({\it top right}
   in units of K and spectra ({\it bottom}) for the
   SiO(2-1) transition for the reference run. 
   The jet is assumed to be in the plane of sky implying an angle of
   inclination of $90^{\circ}$. The vertical {\it magenta dashed} line
   represents the cut for the PV digram and {\it green} crosses marks the positions from which the spectra are taken.} 
\label{empvspec90}
\end{figure}

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig10.pdf}%diffabun_img.pdf}%{\figpath/EM_gaussandth.png}
 \caption{Integrated intensity maps of SiO(2-1) line in units of K-km\,s$^{-1}$, for runs with molecular
   cooling and having different abundance profiles.}
\label{fig:diffabunem}
\end{figure*}

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig13.pdf}%imshkvhcontomaps.png}
 \caption{Contour maps of multi-line integrated SiO
   emission convolved with a 2\arcsec\,beam obtained using parameters
   of the reference run. The line transitions are labelled on the top
   left corner of each panel. The contour colors represent different intensities in K-km\,s$^{-1}$, i.e, 50.0({\it red}), 30.0({\it green}), 20.0({\it blue}), 5.0({\it
     magenta}), 2.0({\it cyan}), 0.1({\it black}).} 
\label{fig:ltconts}
\end{figure*}

\cite{Nisini:2007p13128} have shown that the current
plane-parallel shock models fit reasonably well the physical
conditions traced by the SiO emission i.e., temperature $\lesssim$
1000\,K and H$_{2}$ number density $\sim$ 10$^{5}$-10$^{6}$
cm$^{-3}$, assuming optically thin emission. 
They can also fit the observed fractional SiO abundance of
$\sim$ 10$^{-7}$ in these outflows. However, {\em they fail to predict all
the line profiles and in particular their similarities for low and
high excitation lines, which we are able to obtain.} 
Line emission for SiO from our model also traces similar
physical environment. Fig.~\ref{fig:Hmolform} shows that regions close
to the internal knot have temperatures up to 1000\,K and molecular
hydrogen density of the order of 10$^{6}$ cm$^{-3}$ (also see
fig.~\ref{grid}). These are the regions where the bulk of the SiO
emission is present, with fractional abundances between 10$^{-6}$ to 10$^{-8}$.
The line profiles shapes obtained from our study fit reasonably well that observed by
\cite{Nisini:2007p13128}, especially for the bullets seen in
L1448. {\em Interestingly, in our models, the line profiles are equally broadened for low as well
as high excitation lines}. The top panel of fig.~\ref{fig:lineratio} shows spectra for three line
transitions for the reference run but with an angle of inclination
$i$ = 60$^{\circ}$ and convolved with a single dish beam of
15\arcsec\,to simulate observed integrated
intensities of L1448 bullets observed with large beams from JCMT and
IRAM. Intensities predicted from our reference model, $\int T_{\rm MB}
dV$ = 5.76 K\,km\,s$^{-1}$ for SiO(2-1), 
lie within a factor of two from the observed values (Table 2 of \citealt{Nisini:2007p13128}).
%

@@

\begin{figure*}
 \includegraphics[width=2.\columnwidth]{\figpath/pfig14.pdf}%singledish_new.pdf}%{\figpath/LineRatio_THDeq1.png}
 \caption{{\em{Top}}: Line profiles in SiO (2-1), (5-4) and (8-7) at one
   of the inner knots for the reference molecular cooling run. The profiles are
   obtained when the angle of inclination is 60$^{\circ}$ with respect
 to line of sight. {\em{Bottom left}}: Line intensity ratios SiO(8-7)/(5-4)
and SiO(5-4)/(2-1), as a function of velocity. {\em{Bottom right}}:
Variation of normalized integrated intensity with upper line transition J$_{\rm
  up}$ for the reference run.}
\label{fig:lineratio}
\end{figure*}

\subsection{Slow Component and Grain Chemistry}
%
We have presented a simple model that is able to explain the spectral characteristics
of the SiO EHV component seen
in several young outflows. The reference model
consistently solves for the H$_{2}$ chemistry in the presence of
appropriate cooling. The temperature, density and velocities obtained from dynamical
simulation is used as input in the non-LTE radiative transfer code to obtain SiO maps, spectra and PV
diagrams. One essential ingredient required for the radiative transfer modeling
is the SiO fractional abundance and its dependence on shock velocity. 
Though 1D models that study the formation of SiO in the gas phase from
sputtering and grain-grain 
collisions exists, their focus is mainly on weaker C-type shocks
\citep{Schilke:1997p14140, Caselli:1997p14853, Gusdorf:2008p13800, VanLoo:2013p16573}. 
The pulsating jet propagation model presented here have bulk velocities
reaching up to 120 km\,s$^{-1}$. We have used few
simplified SiO abundance profiles as a function of jet
velocity. Though the shape of the profile is uncertain, the upper and
lower limits used are backed by observational evidences. Ideally, for a consistent
dependence of SiO abundance on shock velocity, one would have to also
solve for silicon and SiO chemistry in a manner similar to that of
H$_{2}$. However, such a numerically expensive model 
with complex chemistry is beyond the scope of this paper 
as this could imply to include grain processing and more gas-phase
chemistry in an MHD multi-fluid approach.  
%

@@

\section{Treatment of Adiabatic Index $\gamma$}
\label{sec:treatgamma}
In dealing with a gas chemistry network which can have both monatomic and diatomic
species, one is immediately presented with the issue of how to treat the adiabatic
index, $\gamma$, of the gas, which directly influences the equation of
state, eq.~\ref{EOS}, and thus the behavior of the gas.
In standard hydrodynamical codes, the gas is generally
assumed to be a monatomic ideal gas, with $\gamma$ = 5/3. 
A diatomic gas, such as H$_{2}$, on the other hand, has a value of $\gamma$
= 1.4 at low temperatures, decreasing to below 1.3
at higher temperatures as the vibrational modes of the hydrogen molecule become
excited. Thus, the value of $\gamma$ for a gas containing a mixture of atomic and
molecular hydrogen requires a temperature-dependent calculation of the $\gamma$
value for H$_2$ and a weighting of the values for the atomic and molecular components
according to their abundance. \cite{Yoshida:2006p13376} have suggested
a prescription of an effective $\gamma$ which takes into account the
corresponding weights of each chemical species and is given by,
\begin{equation}
\frac{1}{\gamma_{\rm eff} - 1} = \frac{1}{\mu}\left[\sum_{i}\frac{X_{i}}{A_{i}}\frac{1}{(\gamma_{i} - 1)}\right]^{-1},
\label{eq:gammaeff}
\end{equation}
where i is the index for each specie and $\mu$ is the mean molecular
weight. The mass fraction is added in the above equation following the
correction suggested in \cite{OSullivan:2009p10078}.
Additionally, the temperature dependence in the case of
H$_{2}$ molecule can be obtained using \cite{landau:1980},
\begin{equation}
\frac{1}{\gamma_{H_{2}} - 1} = \frac{1}{2}\left[ 5 +
  2x^{2}\frac{e^{x}}{(e^{x} - 1)^{2}}\right],
\label{eq:gammaH2}
\end{equation}
where the last term with  x = 6100 $K$/T accounts for the vibrational
degree of freedom, with T being the gas temperature.
%
